Relationships in the extremes and their influence on competition and coexistence

### 2021-08-03 paper/Paper.Rnw
working on storage effect paragraph - more concise

%Old version, prior to my edits: Central to ecology is the maintenance of biodiversity in competitive ecosystems. Populations subject to strong
%competitive pressure may succumb to competitive exclusion - an idea which postulates if species within the same
%trophic level and similar niches exist, one will eventually outcompete the other (Hardin, 1960). Mediation of
%competitive pressure and exclusion allow for species to stably coexist. The criteria for coexistence and competitive
%exclusion can be broken down into contributing mechanisms provided by modern coexistence theory, a framework
%which seeks to assess and quantify mechanisms that result in coexisting species and thus the maintenance of
%biodiversity (Chesson, 1994).
%Now the new, edited version, with some additional changes requested, to be made by Jasmin, please:
The maintenance of biodiversity in competitive ecosystems is a research topic at the absolute core of modern ecology (\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
%***DAN: Jasmin, add a few classic references. See the first sentence of Ellner et al 2019, read the abstracts of those papers
%(more than the abstracts if you want), and consider using them as references here.
The competitive exclusion principle, covered, for instance, in most or all introductory ecology courses and texts, states 
that two species competing for the same limiting resource cannot coexist at constant densities - one will outcompete 
the other and drive it extinction \citep{Hardin_1960}.
Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
framework for understanding how stable species coexistence 
nevertheless occurs, and biodiversity in maintained, in real ecosystems, through stabilizing mechanisms and niche 
differentiation, in spite of the conclusions drawn for idealized scenarios by the competitive exclusion principle. 
MCT breaks down coexistence into the contributions of various mechanisms, with name such as `relative nonlinearity
in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{}. 
%***DAN: Jasmin, fill in Y and Y with names of systems, and accompanying reference(s), 
%in a similar way to how I filled in the Angert example.
%Please also read the abstract of the Angert reference to make sure it is appropriate to cite it in this way (and so that
%you become aware of that reference). One place to look for more references to put in for Y and Z is the top of the 
%right column on the first page of the Barabas review (Ecol Monogr, 2018).
even though it is mathematically complex in the eyes
of most ecologists (but see \cite{Ellner_2016,Ellner_2019} for efforts to make it more readily applied).
%***DAN: Jasmin, note the differences carefully, please, between your version (commented, above) and the edits I've 
%made. It's the same in main structure, making all the same points. It's just a bit more fleshed out. One of the purposes
%of an Intro is to link the reader to the appropriate background lit. You need a fair number of references and a little
%more detail than you previously gave to do that.

@@

other in the left (figure \ref{fig:noise}A, termed ``left-tail association'') or right (figure \ref{fig:noise}A,

@@

%***DAN: Please combine the following two paragraphs into one single paragraph with this topic sentence:
Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence, in particular: storage effects. We review, conceptually, what storage effects are and we explain 
why ATAs may be relevant to them.
%***DAN: I have inserted some comments below that might help.

Storage effect contributes as a stabilizing mechanism to coexistence when there is covariation between competitive pressures and environmental fluctuations.
Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
%***DAN: You don't need to explain why it is reasonable to suspect something, that belabors the point too much. Just say indicate that storage effects
%involve environment-coexistence associations, and hence there is the potential for ATAs to influence storage effects.
Uniquely, storage effect is the only mechanism of modern coexistence theory which depends on the variance between two drivers rather than one.
While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coexistence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms of coexistence.
%***DAN: Now that you have written this, I see it belabors the point too much. It sounds defensive to explain this early on that none of the 
%other mechanisms are influenced. I think instead it's enough to just explain, as laid out in 
%new, proposed topic sentence above, that storage effects are influenced by ATAs. That justifies the study, and then the reader can 
%discover later that only storage effects are influenced. If a referee complains, I think we will be able to effectively respond after the fact.

For clarification, we explain the how storage effects operate in maintaining biodiversity and the conditions that are required for it to do so.
%***DAN: the above sentence will be superceded by the new topic sentence, above, of the combined paragraph, so you can cut this.
We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because.
%***DAN: Protesting too much, sounds defensive. Maybe we can return to this in the Discussion, after we've made our points, but it 
%currently seems to me even that won't be necessary. We can just point out in the first para of the Discussion something like "We have learned
%that ATAs do meaningfully influence species coexistence, and they do so through storage effects." Less defensive. 
Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
In order for storage effect to be a functioning mechanism in coexistence, populations must fluctuate differently in response to environmental fluctuations, there must be a relationship between envrionmental responses and competition, and there must variation in responses at different life history stages. 
In a simple two species system, differences in environmental response allow for one species to grow to high densities while the other is very low. 
Competition is positively related to environmental response for the high density species (species A), stabalizing its growth. 
Conversely, competition is more negatively related to environmental response for the low density species (species B), leading to large fluctuations in population growth rate in both directions. 
Seemingly, this may appear like a scenario for extiction of species B, but it can, in fact, be at an advantage due to differences in life history stages. 
Different responses at different life stages buffer the species against extinction by limiting negative growth rates, but not positive ones. 
Examples of life stages that can buffering against negative growth rates are insensitive adults or dormant seeds that persist even when environmental response amoung others in the population is low \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.  
Thus, higher fluctuation in growth rate can lead to a higher long term mean growth rate. 
In this paper, we investigate the consequences of having ATAs in the relationship between fluctuations environmental response and competitive pressure. More specifically, we answer: Does having higher association in a tail end of the EC relationship, while maintaining overall association, change the contribution of storage effect?     
%***DAN: This explanation of storage effects is too long. See if you can cut it to approximately half its current length.

>>>>

%Old version, prior to my edits: Central to ecology is the maintenance of biodiversity in competitive ecosystems. Populations subject to strong
%competitive pressure may succumb to competitive exclusion - an idea which postulates if species within the same
%trophic level and similar niches exist, one will eventually outcompete the other (Hardin, 1960). Mediation of
%competitive pressure and exclusion allow for species to stably coexist. The criteria for coexistence and competitive
%exclusion can be broken down into contributing mechanisms provided by modern coexistence theory, a framework
%which seeks to assess and quantify mechanisms that result in coexisting species and thus the maintenance of
%biodiversity (Chesson, 1994).
%Now the new, edited version, with some additional changes requested, to be made by Jasmin, please:
The maintenance of biodiversity in competitive ecosystems is a research topic at the absolute core of modern ecology (\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
%***DAN: Jasmin, add a few classic references. See the first sentence of Ellner et al 2019, read the abstracts of those papers
%(more than the abstracts if you want), and consider using them as references here.
The competitive exclusion principle, covered, for instance, in most or all introductory ecology courses and texts, states 
that two species competing for the same limiting resource cannot coexist at constant densities - one will outcompete 
the other and drive it extinction \citep{Hardin_1960}.
Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
framework for understanding how stable species coexistence 
nevertheless occurs, and biodiversity in maintained, in real ecosystems, through stabilizing mechanisms and niche 
differentiation, in spite of the conclusions drawn for idealized scenarios by the competitive exclusion principle. 
MCT breaks down coexistence into the contributions of various mechanisms, with name such as `relative nonlinearity
in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{}, 
%***DAN: Jasmin, fill in Y and Y with names of systems, and accompanying reference(s), 
%in a similar way to how I filled in the Angert example.
%Please also read the abstract of the Angert reference to make sure it is appropriate to cite it in this way (and so that
%you become aware of that reference). One place to look for more references to put in for Y and Z is the top of the 
%right column on the first page of the Barabas review (Ecol Monogr, 2018).
even though it is mathematically complex in the eyes
of most ecologists (but see \cite{Ellner_2016,Ellner_2019} for efforts to make it more readily applied).
%***DAN: Jasmin, note the differences carefully, please, between your version (commented, above) and the edits I've 
%made. It's the same in main structure, making all the same points. It's just a bit more fleshed out. One of the purposes
%of an Intro is to link the reader to the appropriate background lit. You need a fair number of references and a little
%more detail than you previously gave to do that.

@@

other in the left (figure \ref{fig:noise}A, termed ``left-tail association'') or right (figure \ref{fig:noise}C,

@@

%***DAN: Please combine the following two paragraphs into one single paragraph with this topic sentence:
Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence, in particular: storage effects. We review, conceptually, what storage effects are and we explain 
why ATAs may be relevant to them.
%***DAN: I have inserted some comments below that might help.
Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance] \citep{Chesson_1994, Barabas_2018}.
%Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
The realiance of storage effects on the relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects. 
In addition to EC covariance, functioning storage effects also require that conspecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses. 
In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low. 
Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
Conversely, competition is more negatively related to environmental response for the low density species, leading to large fluctuations in population growth rate in both directions. 
Although, this may appear like a scenario for extiction, it can, in fact, be at an advantage due to differences in between life stages, which buffer the species against extinction by limiting negative growth rates, but not positive ones. 
For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overeall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.   
Thus, higher fluctuation in growth rate can lead to a higher long term mean growth rate. 
In this paper, we investigate the consequences of having ATAs in the relationship between fluctuations environmental response and competitive pressure. More specifically, we answer: Does having higher association in a tail end of the EC relationship, while maintaining overall association, change the contribution of storage effect?  
%***DAN: You don't need to explain why it is reasonable to suspect something, that belabors the point too much. Just say indicate that storage effects
%involve environment-coexistence associations, and hence there is the potential for ATAs to influence storage effects.
%Uniquely, storage effect is the only mechanism of modern coexistence theory which depends on the variance between two drivers rather than one.
%While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coexistence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms of coexistence.
%***DAN: Now that you have written this, I see it belabors the point too much. It sounds defensive to explain this early on that none of the 
%other mechanisms are influenced. I think instead it's enough to just explain, as laid out in 
%new, proposed topic sentence above, that storage effects are influenced by ATAs. That justifies the study, and then the reader can 
%discover later that only storage effects are influenced. If a referee complains, I think we will be able to effectively respond after the fact. 
%For clarification, we explain the how storage effects operate in maintaining biodiversity and the conditions that are required for it to do so.
%***DAN: the above sentence will be superceded by the new topic sentence, above, of the combined paragraph, so you can cut this.
%We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because.
%***DAN: Protesting too much, sounds defensive. Maybe we can return to this in the Discussion, after we've made our points, but it 
%currently seems to me even that won't be necessary. We can just point out in the first para of the Discussion something like "We have learned
%that ATAs do meaningfully influence species coexistence, and they do so through storage effects." Less defensive.

%***DAN: This explanation of storage effects is too long. See if you can cut it to approximately half its current length.
